Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-05-25 paper_jasa.tex
comments on p-value
massaged the pvalue explanation, added some questions about Type I, II error

\begin{dfn}\label{dfn:pvalue}
The $p$-value of a lineup of size $m$ is given as 
\[
P(X \ge x) = 1 - B_{K, 1/m} (x-1) = 1- \sum_{i=0}^{x-1} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i},
\]
for $x > 0$, with $X$ defined as above.
\end{dfn}
\noindent When $K=1, m=20$ the $p$-value can only be $1/20=0.05$. When $K=2, m=20$ and if $x=2$ the $p$-value is $0.0025$, but if $x=1$ the $p$-value is$0.905$. When $K=3, m=20$

\begin{table}[htp]
\begin{center}
\begin{tabular}{rrrr}
K & m & x & $p$-value \\\hline
1 & 20 & 1 & 0.0500 \\\hline
2 & 20 & 1 & 0.9050 \\
2 & 20 & 2 & 0.0025 \\\hline
\end{tabular}
\end{center}
\caption{}
\label{pvalue}
\end{table}

@@

The resolution of the Type I error is controlled by both $m$, the size of the lineup, and $K$, the number of observers, with $(1/m)^K$ being the minimal value at which we can control the test. 
%The number of plots in the lineup, $m$, controls the Type I error. The probability of choosing any one of the plots is $1/m$, so the probability of making a wrong decision by chance is $1/m$. 
Classically, Type II error is the probability of failing to reject $H_0$ when it is not true, which is more difficult to quantify.

\begin{dfn}\label{dfn:error}
The \textbf{Type I error} is the probability of \red{rejecting $H_0$ when it is true, the probability for that is $P(X \ge x_{1-\alpha})$, which is controlled by $\alpha$. Note that due to the discrete nature of $X$ we can not always control exactly for $\alpha$. }
% identifying the observed data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots. 
\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, \red{i.e. $P( X \le x_{1-\alpha})$, where $x_{1 - \alpha}$ is the (approximate) $1-\alpha$ quantile of  $B_{K, 1/m}$, the distribution of $X$ under the null hypothesis.}
\red{Under the alternative hypothesis, we have to account for different visual abilities of  individuals: let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise.  Then $X_i \sim B_{1, p_i}$ and $X = \sum_i X_i$ has  a Poisson-Binomial distribution.}
\end{dfn}

@@

\green{Ok, I'm really getting confused now. I think that it might be better to define p-value, and significance level next, and then talk about power. Power and p-value are confounded in the discussion below.}

>>>>

\begin{dfn}\label{dfn:pvalue}
The $p$-value of a lineup of size $m$ is given as 
\[
P(X \ge x) = 1 - \sum_{i=0}^{x-1} B_{K, 1/m} (i) = 1- \sum_{i=0}^{x-1} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i},
\]
for $x > 0$, with $X$ defined as above.
\end{dfn}
\noindent Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m$. It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the observed data plot.

\begin{table}[htp]
\begin{center}
\begin{tabular}{|rrr|rrr|rrr|rrr|rrr|}\hline
$K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$  & $x$ & $p$-value & $K$  & $x$ & $p$-value\\\hline
1 &  1 & 0.0500 & 2 &  1 & 0.9050 & 3 & 1 & 0.1426 & 4 & 1 & 0.1855 & 5 & 1 & 0.2262 \\%\cline{1-3}
&&& 2 &  2 & 0.0025 & 3 & 2 & 0.0073 & 4 & 2 & 0.0140 & 5 & 2 & 0.0226 \\
&&& &&& 3 & 3 & 0.0049 & 4 & 3 & 0.0005 & 5 & 3 & 0.0012 \\%\cline{1-6}
&&&  &    &        &   &   &        & 4 & 4 & 0.0000 & 5 & 4 & 0.0000 \\\hline
\end{tabular}
\end{center}
\caption{Possible $p$-values for different numbers of observers, $K$, for fixed size $m$ lineups.}
\label{pvalue}
\end{table}

@@

Associated with any test is the risk of making a Type I error, or Type II error, which for visual inference are defined as follows:

\begin{dfn}\label{dfn:error}
The \textbf{Type I error} is the probability of rejecting $H_0$ when it is true, the probability for that is $P(X \ge x_{1-\alpha})$, which is controlled by $\alpha$.
% identifying the observed data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots. 
\textbf{Type II error} is the probability of failing to identify the observed data plot, when $H_0$ is not true, $P( X \le x_{1-\alpha})$, where $x_{1 - \alpha}$ is the (approximate) $1-\alpha$ quantile of  $B_{K, 1/m}$.
\end{dfn}

\green{Why are we including $x_{1-\alpha}$ in both sides?}

\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, finding an exact $x_{1-\alpha}$. For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$.

\green{check that we consistently use $1/m$, or $\frac{1}{m}$ in the text}.

%The number of plots in the lineup, $m$, controls the Type I error. The probability of choosing any one of the plots is $1/m$, so the probability of making a wrong decision by chance is $1/m$.

\green{Why???? Under the alternative hypothesis, we have to account for different visual abilities of  individuals:} let $X_i$ be a binary random variable with $X_i = 1$, if individual $i$ identifies the data plot from the lineup, and $X_i = 0$ otherwise.  Then $X_i \sim B_{1, p_i}$ and $X = \sum_i X_i$ has  a Poisson-Binomial distribution.

@@
